### 2023-01-14 preprint.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@248cde172ef469d2562cfb1e543110f64eef9496 🚀
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@fc540949f9b43d58c34c27f827e5607a20741112 🚀
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@7aa656d41d88975532ffd237498f5e8ea2810b9f 🚀
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@c62b1bbd14631093c54d29eb7c03d729d3efe513 🚀

\medskip
\begin{flushleft}
{\small
%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}, %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{3,2}, %
\href{https://orcid.org/0000-0002-6004-4027}{Laura\,Pollock}%
%
\,\textsuperscript{1,2}, %
\href{https://orcid.org/0000-0001-6075-8081}{Andrew\,Gonzalez}%
%
\,\textsuperscript{1,2}
\vskip 1em
\textsuperscript{1}\,McGill University; \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences; \textsuperscript{3}\,Université de
Montréal\\
\vskip 1em
\textbf{Correspondance to:}\\
Michael D. Catchen --- \texttt{michael.catchen@mcgill.ca}\\
}
\end{flushleft}

@@

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes. Still
a robust understanding of (and ability to predict and forecast)
interactions among species remains elusive. This knowledge-gap is
largely driven by a shortfall of data---although species occurrence data
has rapidly increased in the last decade, species interaction data has
not kept pace, largely due to the intrinsic difficulty and effort
required to sample interactions. These sampling challenges bias data and
hinder inferences about the structure and dynamics of interactions
networks. Here, we demonstrate the realized false-negative rate (the
percentage of species that actually interact but for which we do not yet
have a record) can be quite high, even in thoroughly sampled systems,
due to the intrinsic variation in abundances across species in a
community. We illustrate how a null model of occurrence detection can be
used to estimate the false-negative rate in a given dataset. One
hypothesis is that interactions between ``rare'' species are themselves
rare because these species are less likely to encounter one-another than
species of higher relative abundance. However, we demonstrate that
across several datasets of spatial or temporally replicated networks,
there are positive associations that suggest these interactions actually
exist but just are not observed. Finally, we assess how false negatives
influence various models of network prediction, and recommend directly
accounting for observation error in predictive models. We conclude by
discussing how the understanding of false-negatives can inform how we
design monitoring schemes for species interactions.\\%

@@

Here we show how to incorporate imperfect observation (both
false-negatives and false-positives) into model predictions of
interaction probability. Models for interaction prediction typically
yield a probability of interaction between each pair of species. When
these are considered with uncertainty, it is usually model-uncertainty,
e.g.~the variance in the interaction probability prediction across
several cross-validation folds, where the data is split into training
and test sets several times. Here we introduce a method for resampling
interaction probabilities (outlined in figure
fig.~\ref{fig:resampling_concept}) that simulates the observation
process with prior estimates of both false-negative and false-positive
probabilities to produce an output distribution of interaction
probabilities that accounts for observation error. We implement this in
the software package InteractionUncertaintySampler.jl.

\begin{figure}
\hypertarget{fig:resampling_concept}{%
\centering
\includegraphics{./figures/uncertainty_sampler.png}
\caption{(A) The process for estimating the false-negative rate (FNR)
for an interaction dataset consisting of \(N\) total observed
interactions. (B) The method for resampling interaction probability
based on estimates of false-negative and false-positive rates. (C) The
method for interaction probability resampling applied to three mammals
and three parasites from the Hadfield \emph{et al.} (2014)
dataset.}\label{fig:resampling_concept}
}
\end{figure}

We do this by using the process for simulating a distribution of null
false-negative rates for a given dataset as described above
(fig.~\ref{fig:resampling_concept} (A)). We then consider the output
prediction from an arbitrary prediction model, which is the probability
\(p_{ij}\) that two species \(i\) and \(j\) interact. To get an estimate
of \(p_{ij}\) that accounts for observation error, we resample the
probability of each interaction \(p_{ij}\) by simulating a series of
particles. Each particle is the product of the resampling algorithm
(``Resampling'' within fig.~\ref{fig:resampling_concept} (B)), where the
value of a particle is first drawn from a Bernoulli with weight
\(p_{ij}\), and if that value is true, the value remains true with
probability \(1-p_{fn}\), and if the first draw is false, the value
remains false with probability \(1-p_{fp}\). Over many samples of
particles, the resulting frequency of `true' outcomes is a single
resample of the interaction probability \(p_{ij}^*\). Across many
resamples, this forms a distribution of probabilities which are adjusted
by the true and false negative rates.

There is also an analytic way to represent approximate this distribution
using the normal approximation to binomial. As a reminder, as the total
number of samples \(n\) from a binomial distribution with success
probability \(p\) from approaches infinity, the sum of total successes
across all samples approaches a normal distribution with mean \(np\) and
variance \(np(1-p)\). We can use this to correct the estimate \(p_{ij}\)
based on the expected false-negative-rate \(p_{fn}\) and false-positive
rate \(p_{fp}\) to obtain the limiting distribution as the number of
resamples approaches infinity for the resampled \(p_{ij}^*\) for a given
number of particles \(n_p\). We do this by first adjusting for the rates
of observation error to get the mean resampled probability,
\(\bar{p_{ij}^*}\), as

\[\bar{p_{ij}^*} = p_{ij}(1+p_{fp})(1-p_{fn})\]

@@

\begin{equation}\protect\hypertarget{eq:eq1}{}{\mathcal{N}\bigg(\bar{p^*_{ij}},
\sqrt{\frac{\bar{p_{ij}^*}(1-\bar{p_{ij}^*})}{n_p}} \bigg)}\label{eq:eq1}\end{equation}

We can then further truncate to remain on the interval \((0,1)\) (as the
output is a probability, although in practice often the probability mass
outside \((0,1)\) is exteremly low.

Why is this useful? Less computation, number of particles \(n_p\) as a
uncertainty width, nautral analogue to number of observations of
co-occurrence. More about this here.

As an example case study, we use a boosted-regresssion-tree to predict
interactions in a host-parasite network (Hadfield \emph{et al.} 2014)
(with features derived in the same manner as Strydom \emph{et al.}
(2021) derives features on this data) to produce a set of interaction
predictions. We then applied this method to a set of a few resampled
interaction probabilities between mammals and parasite species shown in
figure fig.~\ref{fig:resampling_concept} (C). Here we implement a simple
resampling algorithm, but this could be extended, for example by
adjusting the expected FNR for each pair of species by the relative
abundance of each species. We have implemented a general framework for
this resampling methodology in the Julia package
InteractionUncertaintySampler.jl, which enables flexible choices for
priors on false-negative and false-positive rates.

This method allows us to more effectively represent our uncertainty
about observation error in predictions of species interactions. This
also has implications for what we mean by `uncertainty' in interaction
predictions. A model's prediction can be `uncertain' in two different
ways: the model's predictions may have high variance, or the model's
predictions may be centered around a probability of interaction of
\(0.5\), where we are the most unsure about whether this interaction
exists. Improving the incorporation of different forms of uncertainty in
probabilistic interaction predictions seems a necessary next step toward
understanding what pairs of species we know the least about, in order to
prioritize sampling to provide the most new information possible.

\hypertarget{positive-associations-in-co-occurrence-increase-the-false-negative-rate}{%
\section{Positive associations in co-occurrence increase the
false-negative
rate}\label{positive-associations-in-co-occurrence-increase-the-false-negative-rate}}

The model above doesn't consider the possibility that there are positive
or negative associations which shift the probability of species
cooccurrence together due to their interaction (Cazelles \emph{et al.}
2016). However, here we demonstrate that the probability of observing a
false negative can be higher if there is some positive association in
the occurrence of species \(A\) and \(B\). If we denote the probability
that we observe the co-occurrence of two species \(A\) and \(B\) as
\(P(AB)\) and if there is no association between the marginal
probabilities of observing \(A\) and observing \(B\), denoted \(P(A)\)
and \(P(B)\) respectively, then the probability of observing their
co-occurrence is the product of the marginal probabilities for each
species, \(P(AB) = P(A)P(B)\). In the other case where there is some
positive strength of association between observing both \(A\) and \(B\)
because this interaction is ``important'' for each species, then the
probability of observation both \(A\) and \(B\), \(P(AB)\), is greater
than \(P(A)P(B)\) as \(P(A)\) and \(P(B)\) are not independent and
instead are positively correlated, i.e.~\(P(AB)> P(A)P(B)\). In this
case, the probability of observing a single false-negative in our naive
model from fig.~\ref{fig:geometric}(A) is \(p_{fn}= 1-P(AB)\), which due
to the above inequality implies \(p_{fn}>1-P(A)P(B)\). This indicates an
increasingly greater probability of a false negative as the strength of
association gets stronger, \(P(AB) \to P(AB) \gg P(A)P(B)\). However,
this still does not consider variation in species abundance in space and
time (Poisot \emph{et al.} 2015). If positive or negative associations
between species structure variation in the distribution of \(P(AB)\)
across space/time, then the spatial/temporal biases induced by data
collection would further impact the realized false negative rate, as the
probability of false negative would not be constant for each pair of
species across sites.

To test for these positive associations in data we scoured Mangal for
datasets with many spatial or temporal replicates of the same system.
For each dataset, we compute the marginal probability \(P(A)\) of
occurrence of each species \(A\) across all networks in the dataset. For
each pair of interacting species \(A\) and \(B\), we then compute and
compare the probability of co-occurrence if each species occurs
independently, \(P(A)P(B)\), to the empirical joint probability of
co-occurrence, \(P(AB)\). Following our analysis above, if \(P(AB)\) is
greater than \(P(A)P(B)\), then we expect our neutral estimates of the
FNR above to underestimate the realized FNR. In fig.~\ref{fig:mangal},
we see the difference between \(P(AB)\) and \(P(A)P(B)\) for the seven
suitable datasets with enough spatio-temporal replicates and a shared
taxonomic backbone (meaning all individual networks use common species
identifiers) found on Mangal to perform this analysis. Further details
about each dataset are reported in tbl.~\ref{tbl:id}.

@@

\hypertarget{tbl:id}{}
\begin{longtable}[]{@{}ccccccccc@{}}
\caption{\label{tbl:id}This table describes the datasets used in the
above analysis (Fig 2). The table reports the type of each dataset, the
total number of networks in each dataset \((N)\), the total species
richness in each dataset \((S)\), the connectance of each metaweb (all
interactions across the entire spatial-temporal extent) \((C)\), the
mean species richness across each local network \(\bar{S}\), the mean
connectance of each local network \(\bar{C}\), the mean
\(\beta\)-diversity among overlapping species across all pairs of
network species (\(\bar{\beta}_{OS}\)), and the mean \(\beta\)-diversity
among all species in the metaweb (\(\bar{\beta}_{WN}\)). Both metrics
are computed using KGL \(\beta\)-diversity (Koleff \emph{et al.}
2003)}\tabularnewline
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endfirsthead
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endhead
Kopelke \emph{et al.} (2017) & Food Web & 100 & 98 & 0.037 & 7.87 &
0.142 & 1.383 & 1.972\tabularnewline
Thompson \& Townsend (2000) & Food Web & 18 & 566 & 0.014 & 80.67 &
0.049 & 1.617 & 1.594\tabularnewline
Havens (1992) & Food Web & 50 & 188 & 0.065 & 33.58 & 0.099 & 1.468 &
1.881\tabularnewline
Ponisio \emph{et al.} (2017) & Pollinator & 100 & 226 & 0.079 & 23.0 &
0.056 & 1.436 & 1.870\tabularnewline
Hadfield \emph{et al.} (2014) & Host-Parasite & 51 & 327 & 0.085 & 32.71
& 0.337 & 1.477 & 1.952\tabularnewline
Closs \& Lake (1994) & Food Web & 12 & 61 & 0.14 & 29.09 & 0.080 & 1.736
& 1.864\tabularnewline
CaraDonna \emph{et al.} (2017) & Pollinator & 86 & 122 & 0.18 & 21.42 &
0.312 & 1.527 & 1.907\tabularnewline
\bottomrule
\end{longtable}

@@

Species interactions enable the persistence and functioning of
ecosystems, but our understanding of interactions is limited due to the
intrinsic difficulty of sampling. Here we have provided a null model for
the expected number of false-negatives in an interaction dataset. We
demonstrated that we expect many false-negatives in species interaction
datasets purely due to the intrinsic variation of abundances within a
community. We also, for the first time to our knowledge, measured the
strength of association between co-occurrence and interactions (Cazelles
\emph{et al.} 2016) across many empirical systems, and found that these
positive associations are both very common, and showed algebraically
that they increase the realized FNR. We have also shown that
false-negatives could further impact our ability to both predict
interactions and infer properties of the networks, which highlights the
need for further research into methods for correcting this bias in
existing data.

@@

This work has several practical consequences for the design of
interaction samples. Simulating the process of observation could be a
powerful tool for estimating the sampling effort required by a study
that takes relative abundance into account, and provides a null baseline
for expected FNR. It is necessary to take the size of the species pool
into account when deciding how many total samples is sufficient for an
``acceptable'' FNR (fig.~\ref{fig:geometric}(C \& D)). Further the
spatial and temporal turnover of interactions means any approach to
sampling prioritization must be spatiotemporal. We demonstrated earlier
that observed negatives outside of the range of both species aren't
informative, and therefore using species distribution models could aid
in this spatial prioritization of sampling sites.

We also should address the impact of false-negatives on the inference of
process and causality in community ecology. We demonstrated that in
model food webs, false-negatives do not impact the measure of total
trophic levels until very high FNR (figure fig.~\ref{fig:addedfnr}(C)),
although we cannot generalize this further to other properties. This has
immediate practical concern for how we design what taxa to sample---does
it matter if the sampled network is fully connected? It has been shown
that the stability of subnetworks can be used to infer the stability of
the metaweb paper beyond a threshold of samples (Song \emph{et al.}
2022b). But does this extend to other network properties? And how can we
be sure we are at the threshold at which we can be confident our sample
characterizes the whole system? We suggest that modeling observation
error like we have done here can address these questions and aid in the
design of samples of species interactions. To try and sample to avoid
all false-negatives is a fool's errand. Species ranges overlap to form
mosaics, which themselves are often changing in time. Communities and
networks don't end in space, and the interactions that connect species
on the `periphery' of a given network to species outside the spatial
extent of a given sample will inevitably appear as false-negatives in
practical samples. The goal should instead be to sample a system enough
to have a statistically robust estimate of the current state and
empirical change over time of an ecological community at a given spatial
extent and temporal resolution, and to determine what the sampling
effort required prior to sampling.

Our work highlights the need for a quantitatively robust approach to
sampling design, both for interactions (Jordano 2016) and all other
aspects of biodiversity (Carlson \emph{et al.} 2020). As anthropogenic
forces create rapid shifts in our planet's climate and biosphere, this
is an imperative to maximize the amount of ecological information we get
in our finite samples, and make our inferences and decisions based on
this data as robust as possible. Where we choose to sample, and how
often we choose to sample there, has strong impacts on the inferences we
make from data. Incorporating a better understanding of sampling effort
and bias to the design of biodiversity monitoring systems, and the
inference and predictive models we apply to this data, is imperative in
understanding how biodiversity is changing, and making actionable
forecasts about the future of ecological interactions on our planet.

>>>>

\medskip
\begin{flushleft}
{\small
%
\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%
%
\,\textsuperscript{1,2}, %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{3,2}, %
\href{https://orcid.org/0000-0002-6004-4027}{Laura J.\,Pollock}%
%
\,\textsuperscript{1,2}, %
\href{https://orcid.org/0000-0001-6075-8081}{Andrew\,Gonzalez}%
%
\,\textsuperscript{1,4}
\vskip 1em
\textsuperscript{1}\,McGill University; \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences; \textsuperscript{3}\,Université de
Montréal; \textsuperscript{4}\,Québec Centre for Biodiversity Science\\
\vskip 1em
\textbf{Correspondance to:}\\
Michael D. Catchen --- \texttt{michael.catchen@mcgill.ca}\\
}
\end{flushleft}

@@

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes.
Despite their importance, a robust understanding of (and ability to
predict and forecast) interactions among species remains elusive. This
knowledge-gap is largely driven by a shortfall of data---although
species occurrence data has rapidly increased in the last decade,
species interaction data has not kept pace, largely due to the intrinsic
difficulty and effort required to sample interactions. This means there
are many interactions between species that occur in nature, but we do
not think this interaction occurs because we have no record of it. These
so-called ``false-negatives'' bias data and hinder inference about the
structure and dynamics of interaction networks. Here, we demonstrate the
realized rate of false-negatives in data can be quite high, even in
thoroughly sampled systems, due to the intrinsic variation in abundances
across species in a community. We illustrate how a null model of
occurrence detection can be used to estimate the false-negative rate in
a given dataset. We also show how to directly incorporate uncertainty
due to observation error into model-based predictions of interaction
probabilities between species. One hypothesis is that interactions
between ``rare'' species are themselves rare because these species are
less likely to encounter one-another than species of higher relative
abundance, and that this can (in part) explain the common pattern of
nestedness in bipartite interaction networks. However, we demonstrate
that across several datasets of spatial or temporally replicated
networks, there are positive associations between species co-occurrence
and interactions, which suggests these interactions among ``rare''
species actually exist but simply are not observed. Finally, we assess
how false negatives influence various models of network prediction, and
recommend directly accounting for observation error in predictive
models. We conclude by discussing how the understanding of
false-negatives can inform how we design monitoring schemes for species
interactions.\\%

@@

Here we show how to incorporate uncertainty into model predictions of
interaction probability to account for imperfect observation (both
false-negatives and false-positives). Models for interaction prediction
typically yield a probability of interaction between each pair of
species, \(p_{ij}\). When these are considered with uncertainty, it is
usually model-uncertainty, e.g.~the variance in the interaction
probability prediction across several cross-validation folds, where the
data is split into training and test sets several times. The method we
introduce adjusts the value of a model's predictions to produce a
distribution of interaction probabilities, which are adjusted by a given
false-negative-rate \(p_{fn}\) and false-positive-rate \(p_{fp}\)
(outlined in figure fig.~\ref{fig:resampling_concept}). We describe
first how to sample from this distribution of adjusted interaction
probabilities via simulation, and show that this distribution can be
well-approximated analytically.

\begin{figure}
\hypertarget{fig:resampling_concept}{%
\centering
\includegraphics{./figures/uncertainty_sampler.png}
\caption{(A) The process for estimating the false-negative-rate (FNR)
for an interaction dataset consisting of \(N\) total observed
interactions. (B) The method for resampling interaction probability
based on estimates of false-negative and false-positive rates. (C) The
method for interaction probability resampling applied to three mammals
and three parasites from the Hadfield \emph{et al.} (2014)
dataset.}\label{fig:resampling_concept}
}
\end{figure}

We then consider the output prediction from an arbitrary prediction
model, which is the probability \(p_{ij}\) that two species \(i\) and
\(j\) interact. To get an estimate of \(p_{ij}\) that accounts for
observation error, we resample the probability of each interaction
\(p_{ij}\) by simulating a set of several `particles,' where each
particle is a realization of an interaction occurring (either true or
false with probabilities \(p_{ij}\) and \(1-p_{ij}\) respectively) and
then being correctly observed with probabilities given by \(p_{fp}\) and
\(p_{fn}\) to yield a single boolean outcome for each particle.
(``Resampling'' within fig.~\ref{fig:resampling_concept} (B)). Over many
samples of particles, the resulting frequency of `true' outcomes is a
single resample of the interaction probability \(p_{ij}^*\). Across
several samples each of several particles, this forms a distribution of
probabilities which are adjusted by the true and false negative rates.

There is also an analytic way to approximate this distribution using the
normal approximation to binomial. As a reminder, as the total number of
samples \(n\) from a binomial distribution with success probability
\(p\) from approaches infinity, the sum of total successes across all
samples approaches a normal distribution with mean \(np\) and variance
\(np(1-p)\). We can use this to correct the estimate \(p_{ij}\) based on
the expected false-negative-rate \(p_{fn}\) and false-positive rate
\(p_{fp}\) to obtain the limiting distribution as the number of
resamples approaches infinity for the resampled \(p_{ij}^*\) for a given
number of particles \(n_p\). We do this by first adjusting for the rates
of observation error to get the mean resampled probability,
\(\mathbb{E}[{p_{ij}^*}]\), as

\[
\mathbb{E}[{p_{ij}^*}] = p_{ij}(1-p_{fp})+ (1-p_{ij})p_{fn}
\]

@@

\begin{equation}\protect\hypertarget{eq:eq1}{}{p_{ij}^* \sim \mathcal{N}\bigg( \mathbb{E}[p_{ij}^*] , \sqrt{\frac{\mathbb{E}[p_{ij}^*]
(1-\mathbb{E}[p_{ij}^*] )}{n_p}} \bigg)}\label{eq:eq1}\end{equation}

We can then further truncate to remain on the interval \((0,1)\) (as the
output is a probability, although in practice often the probability mass
outside \((0,1)\) is exteremly low. As an example case study, we use a
boosted-regression-tree to predict interactions in a host-parasite
network (Hadfield \emph{et al.} 2014) (with features derived in the same
manner as Strydom \emph{et al.} (2021) derives features on this data) to
produce a set of interaction predictions. We then applied this method to
a set of a few resampled interaction probabilities between mammals and
parasite species shown in figure fig.~\ref{fig:resampling_concept}(C).

Why is this useful? For one, this analytic method avoids the extra
computation required by simulating samples from this distribution
directly. Further, it enables continuous exanination of the number of
particles \(n_p\) as a uncertainty width. The natural analogue for the
number of particles sampled is the number of observations of
co-occurrence for a given pair of species---the fewer the particles, the
higher the variance of the resulting approximation (see supplemental
figure 1 for an example). Equation (\textbf{ref:eq:eq1?}) is undefined
for 0 particles (i.e.~0 observations co-occurrence), although as \(n_p\)
approaches \(0\) the approxated normal (once truncated) approaches a
uniform distribution on the interval \((0,1)\), the maximum entropy
distribution where we have no information about the possibility of an
interaction.

This also has implications for what we mean by `uncertainty' in
interaction predictions. A model's prediction can be `uncertain' in two
different ways: (1) the model's predictions may have high variance, or
(2) the model's predictions may be centered around a probability of
interaction of \(0.5\), where we are the most unsure about whether this
interaction exists . For non-zero \(p_{fn}\) and \(p_{fp}\), the above
model always moves the mean interaction probability toward 0.5
(supplemental figure 2), siding toward higher uncertainty in the latter
(higher entropy) sense. Improving the incorporation of different forms
of uncertainty in probabilistic interaction predictions seems a
necessary next step toward understanding what pairs of species we know
the least about, in order to prioritize sampling to provide the most new
information possible.

\hypertarget{positive-associations-in-co-occurrence-increase-the-false-negative-rate}{%
\section{Positive associations in co-occurrence increase the
false-negative-rate}\label{positive-associations-in-co-occurrence-increase-the-false-negative-rate}}

The model above doesn't consider the possibility that there are positive
or negative associations which shift the probability of species
cooccurrence together due to their interaction (Cazelles \emph{et al.}
2016). However, here we demonstrate that the probability of having a
false-negative can be higher if there is some positive association in
the occurrence of species \(A\) and \(B\). If we denote the probability
that we observe the co-occurrence of two species \(A\) and \(B\) as
\(P(AB)\) and if there is no association between the marginal
probabilities of observing \(A\) and observing \(B\), denoted \(P(A)\)
and \(P(B)\) respectively, then the probability of observing their
co-occurrence is the product of the marginal probabilities for each
species, \(P(AB) = P(A)P(B)\). In the other case where there is some
positive strength of association between observing both \(A\) and \(B\)
because this interaction is ``important'' for each species, then the
probability of observation both \(A\) and \(B\), \(P(AB)\), is greater
than \(P(A)P(B)\) as \(P(A)\) and \(P(B)\) are not independent and
instead are positively correlated, i.e.~\(P(AB)> P(A)P(B)\). In this
case, the probability of observing a single false-negative in our naive
model from fig.~\ref{fig:geometric}(A) is \(p_{fn}= 1-P(AB)\), which due
to the above inequality implies \(p_{fn}>1-P(A)P(B)\). This indicates an
increasingly greater probability of a false negative as the strength of
association gets stronger, \(P(AB) \to P(AB) \gg P(A)P(B)\). However,
this still does not consider variation in species abundance in space and
time (Poisot \emph{et al.} 2015). If positive or negative associations
between species structure variation in the distribution of \(P(AB)\)
across space/time, then the spatial/temporal biases induced by data
collection would further impact the realized false-negative-rate, as the
probability of false negative would not be constant for each pair of
species across sites.

To test for these positive associations in data we scoured Mangal for
datasets with many spatial or temporal replicates of the same system,
which led the the resulting seven datasets set in figure
fig.~\ref{fig:mangal}. For each dataset, we compute the marginal
probability \(P(A)\) of occurrence of each species \(A\) across all
networks in the dataset. For each pair of interacting species \(A\) and
\(B\), we then compute and compare the probability of co-occurrence if
each species occurs independently, \(P(A)P(B)\), to the empirical joint
probability of co-occurrence, \(P(AB)\). Following our analysis above,
if \(P(AB)\) is greater than \(P(A)P(B)\), then we expect our neutral
estimates of the FNR above to underestimate the realized FNR. In
fig.~\ref{fig:mangal}, we see the difference between \(P(AB)\) and
\(P(A)P(B)\) for the seven suitable datasets with enough spatio-temporal
replicates and a shared taxonomic backbone (meaning all individual
networks use common species identifiers) found on Mangal to perform this
analysis. Further details about each dataset are reported in
tbl.~\ref{tbl:id}.

@@

\hypertarget{tbl:id}{}
\begin{longtable}[]{@{}ccccccccc@{}}
\caption{\label{tbl:id}The datasets used in the above analysis (Fig 2).
The table reports the type of each dataset, the total number of networks
in each dataset \((N)\), the total species richness in each dataset
\((S)\), the connectance of each metaweb (all interactions across the
entire spatial-temporal extent) \((C)\), the mean species richness
across each local network \(\bar{S}\), the mean connectance of each
local network \(\bar{C}\), the mean \(\beta\)-diversity among
overlapping species across all pairs of network species
(\(\bar{\beta}_{OS}\)), and the mean \(\beta\)-diversity among all
species in the metaweb (\(\bar{\beta}_{WN}\)). Both metrics are computed
using KGL \(\beta\)-diversity (Koleff \emph{et al.}
2003)}\tabularnewline
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endfirsthead
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endhead
Kopelke \emph{et al.} (2017) & Food Web & 100 & 98 & 0.037 & 7.87 &
0.142 & 1.383 & 1.972\tabularnewline
Thompson \& Townsend (2000) & Food Web & 18 & 566 & 0.014 & 80.67 &
0.049 & 1.617 & 1.594\tabularnewline
Havens (1992) & Food Web & 50 & 188 & 0.065 & 33.58 & 0.099 & 1.468 &
1.881\tabularnewline
Ponisio \emph{et al.} (2017) & Pollinator & 100 & 226 & 0.079 & 23.0 &
0.056 & 1.436 & 1.870\tabularnewline
Hadfield \emph{et al.} (2014) & Host-Parasite & 51 & 327 & 0.085 & 32.71
& 0.337 & 1.477 & 1.952\tabularnewline
Closs \& Lake (1994) & Food Web & 12 & 61 & 0.14 & 29.09 & 0.080 & 1.736
& 1.864\tabularnewline
CaraDonna \emph{et al.} (2017) & Pollinator & 86 & 122 & 0.18 & 21.42 &
0.312 & 1.527 & 1.907\tabularnewline
\bottomrule
\end{longtable}

@@

Species interactions enable the persistence and functioning of
ecosystems, but our understanding of interactions is limited due to the
intrinsic difficulty of sampling them. Here we have provided a null
model for the expected number of false-negatives in an interaction
dataset. We demonstrated that we expect many false-negatives in species
interaction datasets purely due to the intrinsic variation of abundances
within a community. We also, for the first time to our knowledge,
measured the strength of association between co-occurrence and
interactions (Cazelles \emph{et al.} 2016) across many empirical
systems, and found that these positive associations are both very
common, and showed algebraically that they increase the realized FNR. We
have also shown that false-negatives could further impact our ability to
both predict interactions and infer properties of the networks, which
highlights the need for further research into methods for correcting
this bias in existing data.

@@

This work has several practical consequences for the design of surveys
for species' interactions. Simulating the process of observation could
be a powerful tool for estimating the sampling effort required by a
study that takes relative abundance into account, and provides a null
baseline for expected FNR. It is necessary to take the size of the
species pool into account when deciding how many total samples is
sufficient for an ``acceptable'' FNR (fig.~\ref{fig:geometric}(C \& D)).
Further the spatial and temporal turnover of interactions means any
approach to sampling prioritization must be spatiotemporal. We
demonstrated earlier that observed negatives outside of the range of
both species aren't informative, and therefore using species
distribution models could aid in this spatial prioritization of sampling
sites.

We also should address the impact of false-negatives on the inference of
process and causality in community ecology. We demonstrated that in
model food webs, false-negatives do not impact the measure of total
trophic levels until very high FNR (figure fig.~\ref{fig:addedfnr}(C)),
although we cannot generalize this further to other properties. This has
immediate practical concern for how we design what taxa to sample---does
it matter if the sampled network is fully connected? It has been shown
that the stability of subnetworks can be used to infer the stability of
the metaweb paper beyond a threshold of samples (Song \emph{et al.}
2022b). But does this extend to other network properties? And how can we
be sure we are at the threshold at which we can be confident our sample
characterizes the whole system? We suggest that modeling observation
error as we have done here can address these questions and aid in the
design of samples of species interactions. To try to survey to avoid all
false-negatives is a fool's errand. Species ranges overlap to form
mosaics, which themselves are often changing in time. Communities and
networks don't end in space, and the interactions that connect species
on the `periphery' of a given network to species outside the spatial
extent of a given sample will inevitably appear as false-negatives in
practical samples. The goal should instead be to sample a system enough
to have a statistically robust estimate of the current state and
empirical change over time of an ecological community at a given spatial
extent and temporal resolution, and to determine what the sampling
effort required should be prior to sampling.

Our work highlights the need for a quantitatively robust approach to
sampling design, both for interactions (Jordano 2016) and all other
aspects of biodiversity (Carlson \emph{et al.} 2020). As anthropogenic
forces create rapid shifts in our planet's climate and biosphere, this
is an imperative to maximize the amount of ecological information we get
in our finite samples, and make our inferences and decisions based on
this data as robust as possible. Where we choose to sample, and how
often we choose to sample there, has strong impacts on the inferences we
make from data. Incorporating a better understanding of sampling effort
and bias to the design of biodiversity monitoring systems, and the
inference and predictive models we apply to this data, is imperative in
understanding how biodiversity is changing, and making forecasts that
can guide conservation action.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

AG \& MDC acknowledge the support of the Liber Ero Chair for
Biodiversity conservation and NSERC.
